---
title: "Data preapration"
author: "Juliano Palacios-Abrantes & Sarah M. Roberts"
date: "2022-08-19"
output: html_document
---

```{r setup, include = F}
source(here::here('./Functions/function_list.R'))

load_pkg(c(
  # For grid estimation and species selection
  "spatialEco","tidyverse", "geosphere","raster","units","matrixStats","sf","rmapshaper",
  # For collaborative ease
  "here"
  ))

# For grid e
# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)


```

# Identify grids that are MPAs or surrounding an MPA

## load the data into R 

```{r}
MPA1 <- read_sf(here("./Data/spatial/WDPA_WDOECM_Jul2022_Public_marine_shp/WDPA_WDOECM_Jul2022_Public_marine_shp_0/WDPA_WDOECM_Jul2022_Public_marine_shp-polygons.shp"))
#plot(MPA1)

MPA2 <- read_sf(here("./Data/spatial/WDPA_WDOECM_Jul2022_Public_marine_shp/WDPA_WDOECM_Jul2022_Public_marine_shp_1/WDPA_WDOECM_Jul2022_Public_marine_shp-polygons.shp"))

MPA3 <- read_sf(here("./Data/spatial/WDPA_WDOECM_Jul2022_Public_marine_shp/WDPA_WDOECM_Jul2022_Public_marine_shp_2/WDPA_WDOECM_Jul2022_Public_marine_shp-polygons.shp"))

MPA <- dplyr::bind_rows(list(MPA1, MPA2, MPA3))
crs <- "+init=epsg:4326"

rm(MPA1, MPA2, MPA3)

```

#dissolve so that they are not overlapping 
This is necessary for the area calculation 
Here is where you would select out just the Marine (2 or the marine and coastal, 2 and 1)
You can also select out just the no take here. 
```{r}

MPA_dissolve2 <- ms_simplify(MPA, keep = 0.005,
                                keep_shapes = FALSE)
MPA_dissolve2$ID <- as.character(1:nrow(MPA_dissolve2))

ggplot() + 
  geom_sf(data = MPA_dissolve2, aes(fill = ID))

Mode <- function(x) {
 ux <- unique(x)
 ux[which.max(tabulate(match(x, ux)))]
}

```

##which gridcells overlap with an MPA (and what is the area of that gridcell that contains and MPA)

```{r}
grid <- read_sf(here("./Data/spatial/worldsq_EA.shp"))

#convert grid coordinate system to be same as lat long WGS_1984
grid <- st_transform(grid, crs = st_crs(crs))

#this tells you which gridcells intersect with an MPA and what area of the gridcell is covered 
grid_intersection <- st_intersection(grid, MPA_dissolve2) %>% 
  group_by(Seq) %>% 
  summarise(geom = st_union(geometry), MPA_ID = Mode(ID)) %>%
  mutate(geom = st_sfc(geom),
         area = st_area(geom))
rm(MPA_dissolve2)
grid_intersection$status <- "protected" #lets us know that those grid cells have some level of MPA wiithin them 

#join this back to the big grid (right now we just have the gridIDs that are overlapping an MPA)
grid_intersection_2 <- as.data.frame(grid_intersection)
grid2 <- as.data.frame(grid)
grid2 <- left_join(grid2, grid_intersection_2, by = "Seq")
grid2$protected_area_m2  <- as.numeric(grid2$area)
grid2 <- grid2 %>% dplyr::select(-geom)
grid3 <- grid2
grid2 <- st_as_sf(grid2)




# ggplot() + geom_sf(subset(grid2, grid2$status == "protected"), mapping = aes(fill = protected_area_m2))+ geom_sf(grid_intersection, mapping = aes(colour = "red"))

#that looks like it worked to me. 

#we can join back to the gridPoints file for the next step 
gridpoints <- read.csv(here("./Data/Lon_Lat_DBEM.txt"), header = FALSE)
colnames(gridpoints) <- c("Seq", "long", "lat")
gridpoints <- left_join(gridpoints, grid3, by = "Seq")

# ggplot() +
#   geom_point(data = subset(gridpoints, gridpoints$status == "protected"), aes(x = Lon, y = Lat), size = .2, colour = "red") 

grid_sp <- st_as_sf(gridpoints, coords = c("long", "lat"), crs = st_crs(crs))

```

#figure out surrounding grids 
```{r}
grid_sp_MPA <- grid_sp %>% dplyr::filter(status == "protected")
#st_crs(grid_sp_MPA) <- NA
grid_sp_no_MPA <- grid_sp %>% dplyr::filter(is.na(status))
#st_crs(grid_sp_no_MPA) <- NA

grid_MPA <- gridpoints %>% dplyr::filter(status == "protected")
grid_no_MPA <- gridpoints %>% dplyr::filter(is.na(status))

#within a buffer distance to MPAs

buffer <- st_buffer(grid_sp_MPA,.75)

intr <- st_intersection(grid_sp_no_MPA, buffer, sparse = F)
intr <- intr[colnames(intr) %in% c(colnames(grid_sp_no_MPA), "MPA_ID.1")]
intr$status <- "surrounding"

ggplot() + geom_sf(buffer, mapping = aes(colour = "buffer")) + #geom_sf(subset(grid2, grid2$status == "protected"), mapping = aes(fill = protected_area_m2))+ geom_sf(grid_intersection, mapping = aes(colour = "MPA")) + geom_sf(intr, mapping = aes(colour = "surrounding")) 

dist_touching <-intr
dist_touching <- dist_touching %>% add_count(MPA_ID.1) %>% mutate("prop" = (1/n) + 1) #this is the gridcell with an MPA that they are closest to 

```



#ookay that seemed to work 


#### Clean up the dataset 

```{r}
grid_MPA$prop <- 0
grid_MPA$status <- "protected"

names <- c("Seq", "Lat", "Lon", "area", "status", "prop", "protected_area_m2")

grid_MPA <- grid_MPA[colnames(grid_MPA) %in% names]
dist_touching <- as.data.frame(dist_touching)
dist_touching <- dist_touching[colnames(dist_touching) %in% names]

pt <- rbind(as.data.frame(dist_touching), as.data.frame(grid_MPA))

gridpoints <- gridpoints[colnames(gridpoints) %in% c("Seq","Lon", "Lat")]

grid_final <- left_join(gridpoints, pt, by = c("Seq","Lon", "Lat"))
grid_final <- grid_final %>% distinct() #some of the 

#grid_final %>% dplyr::filter(status %in% c("protected", "surrounding")) %>% ggplot() + geom_point(aes(x = Lon, y = Lat, colour = as.factor(prop)))

write.csv(grid_final, here("./Data/mpa_scenarios/grid_all_mpas.csv"))
```

