---
title: "Data preapration"
author: "Juliano Palacios-Abrantes & Sarah M. Roberts"
date: "2022-08-19"
output: html_document
---

```{r setup, include = F}
source(here::here('./Functions/function_list.R'))

load_pkg(c(
  # For grid estimation and species selection
  "spatialEco","tidyverse", "geosphere","raster","units","matrixStats","sf","rmapshaper", "igraph",
  # For collaborative ease
  "here"
  ))

# For grid e
# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

select <- dplyr::select

```

# Identify grids that are MPAs or surrounding an MPA

## load the data into R 

```{r}
MPA1 <- read_sf(here("./Data/Spatial/WDPA_WDOECM_Jul2022_Public_marine_shp/WDPA_WDOECM_Jul2022_Public_marine_shp_0/WDPA_WDOECM_Jul2022_Public_marine_shp-polygons.shp"))
#plot(MPA1)

MPA2 <- read_sf(here("./Data/Spatial/WDPA_WDOECM_Jul2022_Public_marine_shp/WDPA_WDOECM_Jul2022_Public_marine_shp_1/WDPA_WDOECM_Jul2022_Public_marine_shp-polygons.shp"))

MPA3 <- read_sf(here("./Data/Spatial/WDPA_WDOECM_Jul2022_Public_marine_shp/WDPA_WDOECM_Jul2022_Public_marine_shp_2/WDPA_WDOECM_Jul2022_Public_marine_shp-polygons.shp"))

MPA <- dplyr::bind_rows(list(MPA1, MPA2, MPA3))
crs <- "+init=epsg:4326"


rm(MPA1, MPA2, MPA3)


```

#dissolve so that they are not overlapping 
This is necessary for the area calculation 
Here is where you would select out just the Marine (2 or the marine and coastal, 2 and 1)
You can also select out just the no take here. 
```{r}
beg<- Sys.time()
MPA$to_dissolve <- "A"
MPA_dissolve <- MPA %>% group_by(to_dissolve) %>% summarize()
MPA_dissolve <- st_transform(MPA_dissolve, crs = st_crs(crs))

#ggplot() + 
  #geom_sf(data = MPA_dissolve2, aes(fill = ID))

Mode <- function(x) {
 ux <- unique(x)
 ux[which.max(tabulate(match(x, ux)))]
}

end<- Sys.time()

end-beg
# 2.51 hours

```

##which gridcells overlap with an MPA 
(and what is the area of that gridcell that contains and MPA)

```{r}
grid <- read_sf(here("./Data/spatial/worldsq_EA.shp"))

#convert grid coordinate system to be same as lat long WGS_1984
grid <- st_transform(grid, crs = st_crs(crs)) %>% 
  filter(PWater > 0)

#this tells you which gridcells intersect with an MPA and what area of the gridcell is covered 
grid_intersection <- st_intersection(grid, MPA_dissolve) %>%
  group_by(Seq) %>%
  summarise(geom = st_union(geometry)) %>%
  mutate(geom = st_sfc(geom),
area = st_area(geom))

# write_sf(grid_intersection, here("./Data/grid_intersection.shp"))



#rm(MPA_dissolve2)
grid_intersection$status <- "protected" #lets us know that those grid cells have some level of MPA wiithin them 

#join this back to the big grid (right now we just have the gridIDs that are overlapping an MPA)
grid_intersection_2 <- as.data.frame(grid_intersection)
# write.csv(grid_intersection_2, "grid_intersection_2.csv")

# Load data that takes for ever to make
# grid_intersection_2 <- read_csv(here("Scripts/grid_intersection_2.csv")) %>% 
  # mutate(as.numeric(Seq))
# View(grid_intersection_2)

grid2 <- as.data.frame(grid)
grid2 <- left_join(grid2, grid_intersection_2, by = "Seq")
grid2$protected_area_m2  <- as.numeric(grid2$area)
grid2 <- grid2 %>% dplyr::select(-geom)
grid3 <- grid2
grid2 <- st_as_sf(grid2)

# ggplot() + geom_sf(subset(grid2, grid2$status == "protected"), mapping = aes(fill = protected_area_m2))+ geom_sf(grid_intersection, mapping = aes(colour = "red"))

#that looks like it worked to me. 

#we can join back to the gridPoints file for the next step 
gridpoints <- read.csv(here("./Data/Lon_Lat_DBEM.txt"), header = FALSE)
colnames(gridpoints) <- c("Seq", "Lon", "Lat")
gridpoints <- left_join(gridpoints, grid3, by = c("Seq","Lat","Lon"))

# ggplot() +
#   geom_point(data = subset(gridpoints, gridpoints$status == "protected"), aes(x = Lon, y = Lat), size = .2, colour = "red") 

grid_sp <- st_as_sf(gridpoints, coords = c("long", "lat"), crs = st_crs(crs))

```

#figure out surrounding grids 
```{r}

grid_sp_MPA <- grid_sp %>% dplyr::filter(status == "protected")
#st_crs(grid_sp_MPA) <- NA
grid_sp_no_MPA <- grid_sp %>% dplyr::filter(is.na(status))
#st_crs(grid_sp_no_MPA) <- NA

grid_MPA <- gridpoints %>% dplyr::filter(status == "protected")
grid_no_MPA <- gridpoints %>% dplyr::filter(is.na(status))

```

##proportion calculation 
realocate grid cells that are MPA to the surrounding cells (depending on size of MPA )
Problem is the MPA data has a bunch of overlapping feautres so we dissolved earlier 
now we need to assign every contiguous MPA grid cell a unique group ID 
```{r}


grid_final_sp <- st_as_sf(gridpoints, coords = c("long", "lat"), crs = st_crs(crs))

grid_sp_MPA_poly <- grid_final_sp %>% 
  st_make_grid(cellsize = .5) %>% 
  st_as_sf() %>% 
  st_join(grid_sp)%>% filter(status == "protected") 

# ggplot() + geom_sf(grid_sp_MPA, mapping = aes(colour = "point"))+ geom_sf(grid_sp_MPA_poly, mapping = aes(colour = "grid")) 


touching_list = st_touches(grid_sp_MPA_poly) 

g = graph.adjlist(touching_list)
c = components(g)
grid_sp_MPA_poly$groups = c$membership


# ggplot() + geom_sf(grid_sp_MPA_poly, mapping = aes(colour = as.factor(groups)))

#join that back to the grid_sp_MPA
grid_sp_MPA_poly <- as.data.frame(grid_sp_MPA_poly) %>% dplyr::select(Seq, groups)

grid_sp_MPA <- left_join(grid_sp_MPA, grid_sp_MPA_poly, by = "Seq")

buffer <- st_buffer(grid_sp_MPA,.75)

intr <- st_intersection(grid_sp_no_MPA, buffer, sparse = F)
intr <- intr[colnames(intr) %in% c(colnames(grid_sp_no_MPA), "groups")]
intr$status <- "surrounding"

# ggplot() + geom_sf(buffer, mapping = aes(colour = "buffer")) + geom_sf(subset(grid2, grid2$status == "protected"), mapping = aes(fill = protected_area_m2))+ geom_sf(grid_intersection, mapping = aes(colour = "MPA")) + geom_sf(intr, mapping = aes(colour = as.factor(groups)))


#now do the proportion calculation

grid_sp_MPA <- grid_sp_MPA %>% add_count(groups)
grid_sp_MPA$mpa_n <- grid_sp_MPA$n
grid_sp_MPA1 <- as.data.frame(grid_sp_MPA) %>% dplyr::select(groups, mpa_n) %>% ungroup() %>% distinct()

intr <- left_join(intr, grid_sp_MPA1, by = "groups")

# Proportion is estimated as the number of total protected cells by MPA over the surrounding ones
intr <- intr %>% add_count(groups) %>% mutate("prop" = ((mpa_n/n)/100) + 1) #this is the gridcell with an MPA that they are closest to

# Test plot
# ggplot(intr) +
#   geom_tile(
#     aes(
#       x = Lon,
#       y = Lat,
#       fill = prop,
#       color = prop
#     )
#   )

```

```{r}

write_csv(grid_MPA,"grid_MPA.csv")
write_csv(gridpoints,"gridpoints.csv")


```



#### Clean up the dataset 

```{r}
grid_MPA$prop <- 0
grid_MPA$status <- "protected"

names <- c("Seq", "Lat", "Lon", "area", "status", "prop", "protected_area_m2")

grid_MPA <- grid_MPA[colnames(grid_MPA) %in% names]
intr <- as.data.frame(intr)
# write_csv(intr,"intr.csv")

intr <- intr[colnames(intr) %in% names]

pt <- rbind(as.data.frame(intr), as.data.frame(grid_MPA))
pt <- pt %>% distinct()

# gridpointsb <- gridpoints[colnames(gridpoints) %in% c("Seq","Lon", "Lat")]

gridpoints <- gridpoints %>% 
  dplyr::select(Seq,Lon,Lat)

grid_final <- left_join(gridpoints, pt, by = c("Seq","Lon", "Lat"))
grid_final <- grid_final %>% distinct() #some of the 

grid_final %>% dplyr::filter(status %in% c("protected","surrounding")) %>%
  ggplot() + 
  geom_tile(aes(x = Lon,
                y = Lat,
                colour = prop
                )
            )

# write.csv(grid_final, here("./Data/grid_all_mpas.csv"))
```


### Estimate area


```{r}

 # 1.2 Estimate gridcell area based on global variables (m^2)
# Degree ressolution for area estimation
# https://rdrr.io/github/stineb/rbeni/src/R/area.R
# https://www.pmel.noaa.gov/maillists/tmap/ferret_users/fu_2004/msg00023.html
# https://www.engr.scu.edu/~emaurer/tools/calc_cell_area_cgi.pl

r_earth <- 6371000 # NASA's volumetric mean radious https://nssdc.gsfc.nasa.gov/planetary/factsheet/earthfact.html
res_lon <- 0.5 # lon degree resolution
res_lat <- 0.5 # lat degree resolution

grid_final_area <-
  grid_final %>%
    janitor::clean_names() %>% 
  mutate(
    grid_area = (4 * r_earth^2 * 0.5 * res_lon * pi/180 * cos(abs(lat) * pi/180 ) * sin(0.5 * res_lat * pi/180)),
    prop = ifelse(status == "protected", protected_area_m2/grid_area, prop),
    prop = replace_na(prop,0)
  ) 


write.csv(grid_final_area, here("./Data/grid_all_mpas.csv"))

  ggplot(grid_final_area) +
    geom_histogram(aes(x = prop))
    
grid_final_area %>% 
  ggplot() +
  geom_tile(
    aes(
      x = lon,
      y = lat,
      fill = prop
    )
  )


```

## Prepare grid data for DBEM

In the DBEM, MPA algorithm works as a multiplier of fishing effort so protected areas should = 0 fo fishing effort, regardless of what it is in that cell, is 0. 

```{r}
# Load all MPAs dataset
grid_all_mpas <- read_csv("Data/grid_all_mpas.csv")


mpa_index <- grid_all_mpas %>% 
  # fix numerical mismatch
  mutate(
    status = ifelse(is.na(status),"open",status),
    prop = ifelse(status == "open",1,
                  ifelse(status == "protected" & prop >= 1,0,
                         ifelse(status == "protected" & prop < 1,1-prop,prop)
                  )
    )
  ) %>% 
  select(prop)
  
name <- paste0(here("Data/mpa_scenarios/mpa_index"),".txt")
write.table(mpa_index, file=name, sep="\t", col.names = F, row.names = F)


# mpa_index.txt Sent to Compute Canada on June 2023

```

