---
title: "Data preapration"
author: "Juliano Palacios-Abrantes & Sarah M. Roberts"
date: "2022-08-19"
output: html_document
---

```{r setup, include = F}
source(here::here('./functions/function_list.R'))

load_pkg(c(
  # For grid estimation and species selection
  "spatialEco","tidyverse", "geosphere","raster","units","matrixStats","sf","rmapshaper", "igraph",
  # For collaborative ease
  "here"
  ))

# For grid e
# Fix new updates of sf package
sf::sf_use_s2(use_s2 = FALSE)

select <- dplyr::select


# File path for dealing with multiple computers accessing data folder
if(Sys.info()[8] == "jepa88"){
  root_path <- "~/OneDrive - UBC/Data/mpa_cc_toe/"
}else{
  root_path <- "/Users/sarahroberts/Library/CloudStorage/GoogleDrive-srobs32@gmail.com/My Drive/Work/Projects/mpa_ubc_project/mpa_cc_toe_old" # SARAH MAKE SURE THIS PATH MAKES SENSE
}


```

# Identify grids that are MPAs or surrounding an MPA

## load the data into R 

```{r}
MPA1 <- read_sf(paste0(root_path,"/data/spatial/WDPA_Apr2024_Public_shp/WDPA_Apr2024_Public_shp_0/WDPA_Apr2024_Public_shp-polygons.shp"))
#plot(MPA1)

MPA2 <- read_sf(paste0(root_path,"/data/spatial/WDPA_Apr2024_Public_shp/WDPA_Apr2024_Public_shp_1/WDPA_Apr2024_Public_shp-polygons.shp"))

MPA3 <- read_sf(paste0(root_path,"/data/spatial/WDPA_Apr2024_Public_shp/WDPA_Apr2024_Public_shp_2/WDPA_Apr2024_Public_shp-polygons.shp"))

MPA <- dplyr::bind_rows(list(MPA1, MPA2, MPA3))
#select only marine
MPA <- MPA %>% filter(MARINE == 2)
crs <- "+init=epsg:4326"


rm(MPA1, MPA2, MPA3)

# MPA <- MPA %>% sample_frac(.02) # use this for testing on smaller dataset 

```

#dissolve so that they are not overlapping 
This is necessary for the area calculation 
Here is where you would select out just the no take
MPA <- MPA %>% filter(NO_TAKE == ) "All", "Part", "None", "Not Reported", "Not Applicable"

```{r}
beg<- Sys.time()
MPA$to_dissolve <- "A"
MPA_dissolve <- MPA %>% group_by(to_dissolve) %>% summarize()
MPA_dissolve <- st_transform(MPA_dissolve, crs = st_crs(crs))

#ggplot() + 
  #geom_sf(data = MPA_dissolve2, aes(fill = ID))

Mode <- function(x) {
 ux <- unique(x)
 ux[which.max(tabulate(match(x, ux)))]
}

end<- Sys.time()

end-beg
# 2.51 hours

```

##which gridcells overlap with an MPA 
(and what is the area of that gridcell that contains an MPA)

```{r}
grid <- read_sf(paste0(root_path,"/data/spatial/worldsq_ea/worldsq_EA.shp"))

#convert grid coordinate system to be same as lat long WGS_1984
grid <- st_transform(grid, crs = st_crs(crs)) 

#this tells you which gridcells intersect with an MPA and what area of the gridcell is covered 
grid_intersection <- st_intersection(grid, MPA_dissolve) %>%
  group_by(Seq) %>%
  summarise(geom = st_union(geometry)) %>%
  mutate(geom = st_sfc(geom),
         area = st_area(geom))


grid <- grid 
grid$grid_area <- st_area(grid) #get the area of each grid cell (so we can calculate proportion of mpa)
grid <- grid %>% drop_units()

#rm(MPA_dissolve2)
grid_intersection$status <- "protected" #lets us know that those grid cells have some level of MPA wiithin them 

#join this back to the big grid (right now we just have the gridIDs that are overlapping an MPA)
grid_intersection_2 <- as.data.frame(grid_intersection)
# write.csv(grid_intersection_2, "grid_intersection_2.csv")

# Load data that takes for ever to make
# grid_intersection_2 <- read_csv(here("Scripts/grid_intersection_2.csv")) %>% 
  # mutate(as.numeric(Seq))
# View(grid_intersection_2)

grid2 <- as.data.frame(grid)
grid2 <- left_join(grid2, grid_intersection_2, by = "Seq")
grid2$protected_area_m2  <- as.numeric(grid2$area)


grid2 <- grid2 %>% dplyr::select(-geom)


grid3 <- grid2
grid2 <- st_as_sf(grid2)



# ggplot() + geom_sf(subset(grid2, grid2$status == "protected"), mapping = aes(fill = protected_area_m2))+ geom_sf(grid_intersection, mapping = aes(colour = "red")) +
#     coord_sf(xlim=c(0, 50), ylim=c(40,70), expand = TRUE)

#that looks like it worked to me. 

#we can join back to the gridPoints file for the next step 
gridpoints <- read.csv(paste0(root_path,"/data/spatial/lon_lat_dbem.txt"), header = FALSE)
colnames(gridpoints) <- c("Seq", "Lon", "Lat")
gridpoints <- left_join(gridpoints, grid3, by = c("Seq","Lat","Lon"))

# ggplot() +
#   geom_point(data = subset(gridpoints, gridpoints$status == "protected"), aes(x = Lon, y = Lat), size = .2, colour = "red") 

grid_sp <- st_as_sf(gridpoints, coords = c("Lon", "Lat"), crs = st_crs(crs))

```

#figure out surrounding grids 
```{r}

grid_sp_MPA <- grid_sp %>% dplyr::filter(status == "protected")
#st_crs(grid_sp_MPA) <- NA
grid_sp_no_MPA <- grid_sp %>% dplyr::filter(is.na(status))
#st_crs(grid_sp_no_MPA) <- NA

grid_MPA <- gridpoints %>% dplyr::filter(status == "protected")
grid_no_MPA <- gridpoints %>% dplyr::filter(is.na(status))

```

##proportion calculation 
realocate grid cells that are MPA to the surrounding cells (depending on size of MPA )
Problem is the MPA data has a bunch of overlapping feautres so we dissolved earlier 
now we need to assign every contiguous MPA grid cell a unique group ID 
```{r}


grid_final_sp <- st_as_sf(gridpoints, coords = c("Lon", "Lat"), crs = st_crs(crs))

grid_sp_MPA_poly <- grid_final_sp %>% 
  st_make_grid(cellsize = .5) %>% 
  st_as_sf() %>% 
  st_join(grid_sp)%>% filter(status == "protected") 

# ggplot() + geom_sf(grid_sp_MPA, mapping = aes(colour = "point"))+ geom_sf(grid_sp_MPA_poly, mapping = aes(colour = "grid")) 


touching_list = st_touches(grid_sp_MPA_poly) 

g = graph.adjlist(touching_list)
c = components(g)
grid_sp_MPA_poly$groups = c$membership

cols = rainbow(314, s=.6, v=.9)[sample(1:314,314)] 
ggplot() + geom_sf(grid_sp_MPA_poly, mapping = aes(colour = as.factor(groups)))+ scale_colour_manual(values=cols)+ theme(legend.position="none")+
     coord_sf(xlim=c(0, 50), ylim=c(40,70), expand = TRUE)

#join that back to the grid_sp_MPA
grid_sp_MPA_poly <- as.data.frame(grid_sp_MPA_poly) %>% dplyr::select(Seq, groups)

grid_sp_MPA <- left_join(grid_sp_MPA, grid_sp_MPA_poly, by = "Seq")

buffer <- st_buffer(grid_sp_MPA,.75)

intr <- st_intersection(grid_sp_no_MPA, buffer, sparse = F)
intr <- intr[colnames(intr) %in% c(colnames(grid_sp_no_MPA), "groups")]
intr$status <- "surrounding"

ggplot() + geom_sf(buffer, mapping = aes(colour = "buffer")) + geom_sf(subset(grid2, grid2$status == "protected"), mapping = aes(fill = protected_area_m2))+ geom_sf(grid_intersection, mapping = aes(colour = "MPA")) + geom_sf(intr, mapping = aes(colour = as.factor(groups)))+
     coord_sf(xlim=c(0, 50), ylim=c(40,70), expand = TRUE)


#now do the proportion calculation
#groups is the continuous MPA 

#calculate prop for each MPA 
prop_allocate <- grid_sp_MPA %>% st_drop_geometry()

#protected proportion to reallocate (calculated as area of each mpa/area of grid cells containing mpa)
prop_allocate <- prop_allocate %>% group_by(groups) %>% summarise(prop_allocate = sum(protected_area_m2)/sum(grid_area))

intr <- left_join(intr, prop_allocate, by = "groups")
intr <- intr %>% add_count(groups) %>% mutate("prop" = ((prop_allocate/n)) + 1) #this is the total area of one mpa/total area of grid cells overlapping that MPA/total number of grid cells surrounding that MPA (n) 

```



#### Clean up the dataset 

```{r}
grid_MPA<-grid_MPA %>% st_drop_geometry() %>% mutate(prop = 1- (protected_area_m2/grid_area), status = "protected")

names <- c("Seq", "grid_area", "status", "prop", "protected_area_m2")

grid_MPA <- grid_MPA[colnames(grid_MPA) %in% names]
intr <- intr %>% st_drop_geometry()
# write_csv(intr,"intr.csv")

intr <- intr[colnames(intr) %in% names]

pt <- rbind(as.data.frame(intr), as.data.frame(grid_MPA))
pt <- pt %>% distinct()

# gridpointsb <- gridpoints[colnames(gridpoints) %in% c("Seq","Lon", "Lat")]

gridpoints <- gridpoints %>% 
  dplyr::select(Seq,Lon,Lat)

grid_final <- left_join(gridpoints, pt, by = c("Seq"))
grid_final <- grid_final %>% distinct() #some of the 

grid_final %>% dplyr::filter(status %in% c("protected","surrounding")) %>%
  ggplot() + 
  geom_tile(aes(x = Lon,
                y = Lat,
                fill = prop
                )
            )+
     coord_sf(xlim=c(0, 50), ylim=c(40,70), expand = TRUE)

grid_final <-
  grid_final %>%
    janitor::clean_names() %>% 
  mutate (prop = ifelse(status %in% "protected" | status %in% "surrounding", prop,1))

write.csv(grid_final, paste0(root_path,"/data/mpa/grid_all_mpas.csv"))
```

Sarah stopped here 

prop within an MPA is 1-protected area coverage (so if MPA covers 70% of the grid cell, then prop = 30%, or fishing allowed = 30%)
prop surrounding is 1+ displaced area coverage from MPA (split across all surrounding grid cells)

## Prepare grid data for DBEM

In the DBEM, MPA algorithm works as a multiplier of fishing effort so protected areas should = 0 fo fishing effort, regardless of what it is in that cell, is 0. 


